---
title: "MATH 216 Homework 1"
author: "Brenda Li"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
# For read_csv() command, which is an improved version of base R's read.csv()
library(readr) 
library(knitr)

# Load data sets. Note this assumes this .Rmd files is in the same directory as
# the .csv files.
flights <- read_csv("data/flights.csv") %>% 
  mutate(date=as.Date(date))
weather <- read_csv("data/weather.csv") %>% 
  mutate(date=as.Date(date))
planes <- read_csv("data/planes.csv")
airports <- read_csv("data/airports.csv")
states <- read_csv("data/states.csv")
```





## Admistrative:

Please indicate

* Who you collaborated with: Shannia Fu
* Roughly how much time you spent on this HW so far: 4 Hours
* The URL of the RPubs published URL [here](http://rpubs.com/brendali121/HW1).
* What gave you the most trouble: Determining how to convey all the relevant information in a single plot/table.
* Any comments you have: For question 2, I wasn't sure how to weight the proportion of older planes a carrier used against the number of older planes a carrier used so instead of outputting a single plot, I outputted multiple tables and plots to convey all the relevant information.





## Question 1:

Plot a "time series" of the proportion of flights that were delayed by > 30 minutes on each day.  i.e.
 
* the x-axis should be some notion of time
* the y-axis should be the proportion.

Using this plot, indicate describe the
[seasonality](https://en.wikipedia.org/wiki/Seasonality) of when delays over 30
minutes tend to occur.

```{r, echo=FALSE, message=FALSE, fig.width=12, fig.height=6}

new_flights<-flights %>% 
  mutate(delay_30=dep_delay>30) %>%  
  filter(!is.na(dep_delay)) %>% 
  group_by(date) %>% 
  summarise(p=mean(delay_30)) %>% 
  ggplot(aes(x=date,y=p))+labs(title = "Proportion of Flights Delayed More than 30 Minutes",x="Date",y="Proportion")+geom_point()+geom_smooth(se=FALSE)
new_flights

```

Delays of 30 minutes or longer seem to occur in the summer months of June and July, as well as in the winter months of December and January. This trend could be explained by the fact that these months are the months with the greatest number of travelers, and a higher number of travelers correlates with a greater number of travel delays.    



## Question 2:

Some people prefer flying on older planes.  Even though they aren't as nice,
they tend to have more room.  Which airlines should these people favor?

```{r, echo=FALSE, fig.width=12, fig.height=6}
q2<-inner_join(flights,planes,by="plane") %>% 
  select(plane,year,carrier) %>%
  filter(!is.na(year)) %>% 
  mutate(old=year<2000)

a<-group_by(q2,carrier) %>% 
  summarise(p=round(mean(old),2)) %>% 
  arrange(desc(p)) 
kable(head(a),col.names=c("Carrier", "Proportion of Older Planes"))

b<-group_by(q2,carrier) %>% 
  summarise(num_old=round((mean(old))*n())) %>% 
  arrange(desc(num_old))
kable(head(b),col.names=c("Carrier","Number of Old Planes"))

group_by(q2,year,carrier) %>% 
  summarise(n=n()) %>% 
  ggplot(aes(x=year,y=n))+geom_point()+facet_wrap(~carrier,ncol=5)+labs(title="Distribution of the Years Each Carrier's Planes Were Built",x="Year",y="Number of Planes")

```
The people who prefer flying on older planes should choose to fly on WN since they use over 25000 planes that were built before the year 2000, which is also more than 50% of their planes. You can also note this tendency in the plot of the distributions of the years each carrier's planes were built; you can see that the median of the distribution for WN planes is centered left of the year 2000 and the plot shows a much greater number of planes relative to than all the other airlines whose distributions are also be centered distinctly left of 2000.


## Question 3:

* What states did Southwest Airlines' **flight paths** tend to fly to?
* What states did Southwest Airlines' **flights** tend to fly to?

For example, Southwest Airlines Flight 60 to Dallas consists of a single flight
path, but since it flew 299 times in 2013, it would be counted as 299 flights.

```{r, echo=FALSE, message=FALSE, fig.width=12, fig.height=6}

q3<-inner_join(airports,flights,by=c("iata"="dest")) %>% 
    filter(carrier=="WN") %>% 
  select(state,flight) 

#Flight Paths
a<-q3 %>% 
  group_by(flight,state) %>% 
  tally() %>% 
  group_by(state) %>% 
  summarise(n=n()) %>% 
  arrange(desc(n))
head(a) %>% 
kable(col.names=c("State","Number of Flight Paths"))

#Flights
b<-group_by(q3,state) %>% 
  tally() %>%
  arrange(desc(n))
head(b) %>% 
kable(col.names=c("State","Number of Flights"))

```

Both Southwest flight paths and flights tended to fly into Texas, Florida, Lousiana, California, Oklahoma, and Illonois.




## Question 4:

I want to know proportionately what regions (NE, south, west, midwest) each 
carrier flies to/from Houston in the month of July.  Consider the `month()`
function from the `lubridate` package.

```{r, echo=FALSE, message=FALSE, fig.width=12, fig.height=6}
library(lubridate)
q3<-inner_join(airports,flights,by=c("iata"="dest"))
july<-filter(q3,month(date)==7)

q4<-inner_join(july,states,by="state") %>% 
  select(flight,carrier,region)

flights_per_region <- q4 %>%
  group_by(carrier,region) %>%
  tally() %>%
  rename(num_in_region = n)

flights_per_airline <- q4 %>% 
  group_by(carrier) %>% 
  tally() %>% 
  rename(num_per_carrier=n)

p<-left_join(flights_per_region,flights_per_airline,by="carrier") %>% 
  mutate(prop=num_in_region/num_per_carrier) 

ggplot(data=p,aes(x=carrier,y=prop,fill=region))+geom_bar(stat="identity")+scale_fill_brewer(palette = "YlGnBu")+labs(title="Proportion of Flights that Flies to Each Region",x="Carrier",y="Proportion")

```
